\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-08-31 notes/2loopNLSM.tex
Added first gen unit example.

\def\tree{{\rm tree}}
\def\pol{\varepsilon}
\def\Tr{\, {\rm Tr}}
\def\tr{\, {\rm tr}}
\def\eps{\varepsilon}
\def\e{\varepsilon}
\def\ep{\varepsilon}
\def\SYM{MSYM}
\def\nn{\nonumber}
\def\Neqfour{{{\cal N}=4}}
\def\NeqFour{{{\cal N}=4}}
\def\Neqeight{{{\cal N}=8}}
\def\NeqEight{{{\cal N}=8}}
\def\NeqOne{{{\cal N}=1}}
\def\Fact{{\cal F}}
\def\f{\widetilde f}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\bea{\begin{eqnarray}}
\def\eea{\end{eqnarray}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}
\def\Ksl{{\s K}}
\def\ksl{\s{k}}
\def\Perm{{\cal P}}
\def\M{{\cal M}}
\def\ve{\varepsilon}
\def\tlambda{{\tilde\lambda}}
\def\MHVbar{$\overline{\hbox{MHV}}$}
\def\NMHVbar{$\overline{\hbox{NMHV}}$}
\def\P{{\rm P}}
\def\NHP{{\rm NHP}}
\def\mud{\lambda}
\def\bowtie{{\rm bow\mbox{-}tie}}
\newcommand{\cred}{\bf \color{red}}
 \newcommand{\cblue}{\color{blue}}
 \definecolor{MattOrange}{rgb}{1.0,0.4,0.2}
\newcommand{\cob}{ \bf \color{MattOrange}}
\newcommand{\andd}{\ , \quad \text{and}  \quad}
\newcommand{\forr}{\ , \quad \text{for}  \quad}

@@

\newcommand{\afour}{\ensuremath{A_4^{\text{tree}}}}
\definecolor{NUpurple}{RGB}{078,042,132}

@@

\paragraph{Unitarity}
\label{sec:unitarity}

The last key constraint is that the ansatz have the correct unitarity cuts, that is, that the integrand corresponds to the correct theory.
Generalized unitarity has been reviewed in many places (cite...) so we will not belabor the details here, in particular see (emu) for a recent treatment of unitarity methods for even multiplicity theories like NLSM.
The essence of generalized unitarity is that when internal lines are placed on-shell, the integrand should factorize into a product of on-shell tree amplitudes.
The cut to be performed can, itself, be encoded in a graph.
For example, potential 4pt 2-loop cuts include
\be
\label{eq:PotentialCuts}
\text{(a)} \SubtleCut ~~~~ \text{(b)} \PhysicalCutOne{}{}{}{} ~~~~ \text{(c)} \PhysicalCutTwo{}{}{}{} ~~~~ \text{(d)}  \MaxCut
\ee
as well as many more.
In these graphs, all exposed propagators are understood to be cut/on-shell.
As an equation, a unitarity cut looks schematically like
\be
\label{eq:SchematicCutEquation}
\sum \limits_\text{topologies} \frac{n \vert_\text{cut}}{\prod \limits_\text{uncut} k^2} = \sum \limits_\text{states} \prod \limits_\text{blobs} A_\text{tree} .
\ee
The right hand side of the equation is a product of on-shell tree amplitudes, represented graphically in \cref{eq:PotentialCuts} by grey blobs.
States crossing (or propagating across) the cut must be summed over.
For scalars this sum is trivial but for spinning particles this is no longer the case and the physical state projector must be inserted (see Dimitrius paper for simplifying this operation).
The graph topologies appearing in the left hand side of \cref{eq:SchematicCutEquation} come from blowing up the grey blobs into cubic tree graphs in every conceivable way.
The numerator of each graph is related back to the basis numerators via Jacobi relations.
For each of these topologies, the numerator of the associated graph is subjected to the cut constraints and is then divided by the uncut propagators.
The right hand side is assumed to be known, which places a constraint on the numerator ansatz in the left hand side.

@@

\begin{figure}
  \foreach \x in {1,...,8}
  {
    \begin{minipage}{0.32\textwidth}
      \includegraphics[width=\textwidth]{temp-figs/nonpath-nmc-\x.pdf}
    \end{minipage}
  }
  \caption{The 8 non-pathological next-to-maximal-cut diagrams}
  \label{fig:ym-nmc}
\end{figure}

\ace{\sout{I need to check this.  For now it is a standing
    hypothesis:} Checked.  The failure works as stated.} In fact, we can make the
failure extremely precise.  Start with the diagrams shown in the
Jacobi relation from \cref{eq:TwoLoopJacobiOne}: the double-box,
crossed-box, and penta-triangle.  For each of the three diagrams,
write down the most generic parity-even local ansatz involving four
polarizations and six momenta (any combination of external or loop).
Each ansatz will be 10,010 terms initially.  Next impose the symmetry
constraints on each of the three diagrams, leaving 2761 free
parameters on the double-box, 2576 free parameters on the crossed-box,
and 5040 free paramters on the penta-triangle.  Then impose the
``kissing triangle'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only recieves contributions from the
double-box
\begin{equation}
  \KissingTriangles
  =
  \frac{1}{(\textcolor{nhpBlue}{\ell_1})^2} \
  n \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{(\textcolor{nhpBlue}{\ell_2})^2} \
  n\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied:
\begin{equation}
  \KissingTriangles
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang--Mills representation: locality,
symmetry, and the ``kissing triangle'' cut are incompatible with the
kinematic Jacobi relations.

>>>>

\def\tree{{\rm tree}}
\def\pol{\varepsilon}
\def\Tr{\, {\rm Tr}}
\def\tr{\, {\rm tr}}
\def\eps{\varepsilon}
\def\e{\varepsilon}
\def\ep{\varepsilon}
\def\SYM{MSYM}
\def\nn{\nonumber}
\def\Neqfour{{{\cal N}=4}}
\def\NeqFour{{{\cal N}=4}}
\def\Neqeight{{{\cal N}=8}}
\def\NeqEight{{{\cal N}=8}}
\def\NeqOne{{{\cal N}=1}}
\def\Fact{{\cal F}}
\def\f{\widetilde f}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\bea{\begin{eqnarray}}
\def\eea{\end{eqnarray}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}
\def\Ksl{{\s K}}
\def\ksl{\s{k}}
\def\Perm{{\cal P}}
\def\M{{\cal M}}
\def\ve{\varepsilon}
\def\tlambda{{\tilde\lambda}}
\def\MHVbar{$\overline{\hbox{MHV}}$}
\def\NMHVbar{$\overline{\hbox{NMHV}}$}
\def\P{{\rm P}}
\def\NHP{{\rm NHP}}
\def\mud{\lambda}
\def\bowtie{{\rm bow\mbox{-}tie}}
\newcommand{\cred}{\bf \color{red}}
 \newcommand{\cblue}{\color{blue}}
\newcommand{\andd}{\ , \quad \text{and}  \quad}
\newcommand{\forr}{\ , \quad \text{for}  \quad}

@@

\newcommand{\atree}{\ensuremath{A^{\text{tree}}}}
\newcommand{\afour}{\ensuremath{A_4^{\text{tree}}}}
\definecolor{NUpurple}{RGB}{078,042,132}

@@

\paragraph{Unitarity}
\label{sec:unitarity}
The last key source of constraints comes from ensuring that the ansatz
correctly reproduces the generalized unitarity cuts of the theory.
Matching the unitarity cuts provides the vast majority of
theory-dependent information for the ansatz by ensuring that the
numerators under construction are actually a valid representation of
the theory being considered.  Generalized unitarity methods has been
reviewed in many places \cite{BCFUnitarity, BernHuangReview,
  Bern:2015ooa, Bourjaily:2017wjl, JJHenrikReview, FivePointN4BCJ,
  Bern:2012uf, BCJreview, Edison:2022jln} \ace{Here is a semi-random
  collection of unitarity-related citations.  Probably want to bulk it
  up a bit more} so we will be light on details here. We also want to
call special attention to Ref.~\cite{Carrasco:2023qgz} for a recent
treatment of unitarity methods optimized for even multiplicity
theories like NLSM.

In the context of loop integrand construction, generalized unitarity
allows us to equate products of on-shell tree amplitudes encoded via
graphs, \eg
\begin{equation}
  \label{eq:example-cut-prod}
  \ExCutLabeled{1}{2}{3}{4} = \!\!\!\!\! \sum_{\substack{\text{states of }\\ \{\ell_1, \ell_2, \ell_3, \ell_4\}}} \!\!\!\!\!\!\!
  \atree(1,2,\ell_1,\ell_2) \atree(-\ell_1,-\ell_2,\ell_3,\ell_4) \atree(-\ell_3,-\ell_4,3,4) \,,
\end{equation}
to sums over compatible\footnote{A diagram $g$ is compatible with the
  cut diagram $\gamma$ if $g$ is isomorphic to $\gamma$ or to any of
  the additional factorization channels of $\gamma$. } diagram
numerators evaluated on the support of the on-shell conditions and
weighted by their uncut propagators, \eg
\begin{align}
  \label{eq:example-cut-sum}
  \ExCutLabeled{1}{2}{3}{4} &=
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}{(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7})^2}
  \ n\left[
    \IUCContrib{1}{4}{3}{2}
                              \right] \notag \\
  &\quad + 
  \frac{\delta(\textcolor{cut-color}{\ell_1^2})
    \delta(\textcolor{cut-color}{\ell_2^2})
    \delta(\textcolor{cut-color}{\ell_3^2})
    \delta(\textcolor{cut-color}{\ell_4^2})}
    {(\textcolor{ucp-color}{\ell_5})^2
    (\textcolor{ucp-color}{\ell_6})^2
    (\textcolor{ucp-color}{\ell_7'})^2}
  \ n\left[
    \IUCContrib{1}{3}{4}{2}
    \right] + \text{others}
\end{align}
When drawing cut diagrams, we will use gray blobs as in the left-hand
sides of \cref{eq:example-cut-prod,eq:example-cut-sum} and all drawn
legs are assumed to be on-shell.  When drawing numerator contributions
to cuts, we will not use blobs on vertices, and will use
\textcolor{cut-color}{dashed lines} bisecting edges to denote cut
lines and highlight the \textcolor{ucp-color}{uncut edges}.  From now
on we will leave the cut $\delta(\ell^2)$s implicit.

The essence of generalized unitarity is that when internal lines are
placed on-shell, the integrand should factorize into a product of
on-shell tree amplitudes.  The cut to be performed can, itself, be
encoded in a graph.  For example, potential 4pt 2-loop cuts include
\begin{equation}
\label{eq:PotentialCuts}
\text{(a)} \SubtleCut ~~~~ \text{(b)} \PhysicalCutOne{}{}{}{} ~~~~ \text{(c)} \PhysicalCutTwo{}{}{}{} ~~~~ \text{(d)}  \MaxCut
\end{equation}
as well as many more.
In these graphs, all exposed propagators are understood to be cut/on-shell.
As an equation, a unitarity cut looks schematically like
\begin{equation}
\label{eq:SchematicCutEquation}
\sum \limits_\text{topologies} \frac{n \vert_\text{cut}}{\prod \limits_\text{uncut} k^2} = \sum \limits_\text{states} \prod \limits_\text{blobs} A_\text{tree} \, .
\end{equation}
The right hand side of the equation is a product of on-shell tree
amplitudes, represented graphically in \cref{eq:PotentialCuts} by grey
blobs.  States crossing (or propagating across) the cut must be summed
over.  For scalars this sum is trivial but for spinning particles this
is no longer the case and the physical state projector must be
inserted (see Dimitrius paper for simplifying this operation).  The
graph topologies appearing in the left hand side of
\cref{eq:SchematicCutEquation} come from blowing up the grey blobs
into cubic tree graphs in every conceivable way.  The numerator of
each graph is related back to the basis numerators via Jacobi
relations.  For each of these topologies, the numerator of the
associated graph is subjected to the cut constraints and is then
divided by the uncut propagators.  The right hand side is assumed to
be known, which places a constraint on the numerator ansatz in the
left hand side.

@@

\begin{figure}
  \foreach \x in {1,...,8}
  {
    \begin{minipage}{0.32\textwidth}
      \includegraphics[width=\textwidth]{temp-figs/nonpath-nmc-\x.pdf}
    \end{minipage}
  }
  \caption{The 8 non-pathological next-to-maximal-cut diagrams.
    \ace{Note that Mathematica drew the bottom-left diagram with two
      legs stacked on top of each other which eventually meet in a
      3pt.  One leg is on the ``bottom level'' and connects with a 3pt
      to the two legs coming off to the side, and the other is on the
      ``top level'' and connects to the 4pt.}}
  \label{fig:ym-nmc}
\end{figure}

\ace{\sout{I need to check this.  For now it is a standing
    hypothesis:} Checked.  The failure works as stated.} In fact, we can make the
failure extremely precise.  Start with the diagrams shown in the
Jacobi relation from \cref{eq:TwoLoopJacobiOne}: the double-box,
crossed-box, and penta-triangle.  For each of the three diagrams,
write down the most generic parity-even local ansatz involving four
polarizations and six momenta (any combination of external or loop).
Each ansatz will be 10,010 terms initially.  Next impose the symmetry
constraints on each of the three diagrams, leaving 2761 free
parameters on the double-box, 2576 free parameters on the crossed-box,
and 5040 free paramters on the penta-triangle.  Then impose the
``kissing triangle'' next-to-maximal cut, which in the
\emph{non-planar} $t$-$u$ color channel only recieves contributions from the
double-box
\begin{equation}
  \KissingTriangles
  =
  \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
  n \left[
    \KTDBContrib{1}{2}{3}{4}
  \right]
  +
  \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
  n\left[\KTDBContrib{1}{3}{2}{4}\right] \,.
\end{equation}
After imposing this cut, we find that the original Jacobi relation,
\cref{eq:TwoLoopJacobiOne}, can no longer be satisfied:
\begin{equation}
  \KissingTriangles
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
Thus we have identified the minimal failure state for a
globally-color-dual two-loop Yang--Mills representation: locality,
symmetry, and the ``kissing triangle'' cut are incompatible with the
kinematic Jacobi relations.
